Storm surges in the Strait of Georgia simulated with a regional model

### 2015-06-08 AO_guides/manuscriptAO.tex
Finished Keith's wordy comments

\begin{table}[h]
\centering 
\tbl{{\color{red}Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast. Averages obtained over the surge events discussed on this article. A negative delay indicates that the peak surge occurred later than at Tofino.}} 
{\begin{tabular}{|l |c | c|}
\hline 
Location & Mean surge amplitude (m) & Mean delay from Tofino (hrs) \\
\hline
Port Hardy & 0.491 & -2.67\\
Winter Harbour & 0.572 & 0.333 \\
Tofino  & 0.707 & -- \\
Bamfield & 0.658 & -0.5 \\
Neah Bay & 0.658 & -1.83 \\
Toke Point & 1.17 & -2.83 \\
\hline
\end{tabular}}
\label{tab:pacificsurge}
\end{table}

@@

The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary. The barotropic velocity component of the remote forcing was set to zero. At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).  This scheme is not part of core NEMO 3.4 (the version used here).  Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain. {\color{red} No boundary conditions were set on the baroclinic velocities at the north.}

At coastal boundaries, the partial slip boundary condition, an approximation to no slip, is used. Partial slip allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions. A lateral eddy viscosity of 20 m$^2$ s$^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m$^2$ s$^{-1}$ is used.  Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$. Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ m$^2$ s$^{-1}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively. Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.

@@

In general, the K$_1$ tidal harmonics are well-represented by the model with complex differences below 5 cm at all stations except Seymour Narrows (station 27). Seymour Narrows is a very constricted region in Johnstone Strait and is not well resolved by the model. As such, the M$_2$ amplitude and phase are also poorly reproduced here.  The model's M$_2$ amphidrome, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region. {\color{red} The model also represents the dip in M$_2$ amplitude around stations 25 and 26 reasonably well.} Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.  The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.  {\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised, particularly noticeable at the station closest to the open boundary (Port Renfrew, 1). } Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.

@@

Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range. {\color{red} As such, model water level predictions were corrected by accounting for this error. Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations. As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.  The difference between these two predictions is added to the model sea surface height as a correction. Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing. That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed. The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals. All times are reported in UTC.

@@

The model's skill is reported through a few statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
\begin{equation}
\bar{e} = \text{mean}\left(\left| M - O \right|\right),
\end{equation}
where $M$ and $O$ are the modelled and observed quantities of interest, is a common measure that estimates the average error in the model. Next, a measure comparing the variance of the error to the variance of the observations defined as \citep{thompson2003prediction},
\begin{equation}
\gamma^2 = \frac{\text{var}\left(M-O\right)}{\text{var}\left(O\right)},
\end{equation}
is also calculated. Smaller values of $\gamma^2$ indicate better model skill. This quantity was used by \citet{bernier2006predicting} and \citet{bernier2010tide} in storm surge predictions in Eastern Canada and is calculated here as a point of comparison. A third measure, called the Willmott skill score \citep{willmott1982some}, is also used to assess overall model skill in predicting the water level and residual. This quantity is defined as
\begin{equation}
WS = 1 - \frac{\sum_{i=1}^N \left(M_i - O_i\right)^2}{\sum_{i=1}^N \left(|M_i'| + |O_i'|\right)^2},
\end{equation} 
where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations. This score is bounded by zero and one and values close to one indicate good model skill. Since the Willmott skill score is nondimensional, it is an appropriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals. Since the total water level has a larger variability than the residuals, the $\gamma^2$ measure should be much lower for the total water levels; hence, the Willmott skill score provides an additional point of comparison. These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.  {\color{red} Observed and modelled maximum water levels and surge values as well as differences in their timing are reported in Table \ref{tab:peak}.}

@@

{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and was used to briefly examine stratified effects. When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm. The other stations were not strongly affected. This was also observed in another hindcast (Dec 2006). The effect of stratification was examined by considering a case with an initially horizontally uniform stratification and comparing to the fully stratified case. The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period. Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic model as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}. We plan to investigate stratified effects in more detail in a future study.}

@@

A simple steady state argument balancing the wind stress with the pressure gradient can be used to calculate the slope of the sea surface in a body of water with a constant depth $D$ \citep{pugh2004changing}: 
\begin{equation}
\text{slope} = \frac{C\rho_{air}U^2}{g\rho D},
\end{equation}
where $C$ is the coefficient of friction, $\rho_{air}$ is the density of air, $g$ is the acceleration due to gravity, $U$ is the wind speed, and $\rho$ is the density of water. Given  $C=10^{-3}$, $\rho_{air}=1.23$ kg m$^{-3}$, $\rho= 1035$  kg m$^{-3}$, wind speed $U=20$ m s$^{-1}$, and depth $D=100$ m, the slope of the sea surface is approximately $5\times10^{-7}$. At Point Atkinson, a westerly wind can act over a small distance of approximately 50 km, giving a sea surface elevation on the order of centimetres in agreement with the model findings. Since the Strait of Georgia is not a constant depth and is up to 400 m deep, this calculation is only an approximation. Choosing a larger depth would result in an even smaller estimate for the sea surface elevation caused by wind stresses. Further, northwesterly winds would induce the largest elevations at Point Atkinson due to the large distance over which the winds can act.

%local pressure
Next, the surge amplitude decreases by a few centimetres at all four locations if local atmospheric pressure is not included in the surface forcing. This drop is smaller than the IB sea level change calculated by equation (\ref{eq:inverse}), approximately 1 cm rise in sea level associated with 1 hPa drop in atmospheric pressure. In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level. This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude. In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing. Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response. {\color{red} It is the atmopsheric pressure gradient that drives changes in the sea surface height. Since our domain is relatively small, so is the atmospheric pressure gradient, and hence, we see only a small contribution from local atmospheric pressure. }

@@

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above {\color{red}0.89} and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}). The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals. Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain). {\color{red} Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}

@@

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.  {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia. Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case. However, this is still less than 10 \% of the total surge amplitude. }

@@

{\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.  Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and direction {\color{red} the wind is heading} at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison. Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations. At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012. At Sandheads, both the low resolution and high resolution models have good agreement with observations. The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water. Although not shown here, both models have a similar sea level atmospheric pressure at these locations. Note that data assimilation is used in the parent model of the nested HRDPS product.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
\caption{{\color{red}Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (top), Sandheads (middle) and NOAA Buoy 46087 (bottom). Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey. Observations from Sandheads and Point Atkinson are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.  Observations at the NOAA Buoy are found at this url: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}. Wind direction represents the direction that the wind is headed.} }
\label{fig:dec2012_weather}
\end{figure}

>>>>

\begin{table}[h]
\centering 
\tbl{{\color{red}Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast. Averages obtained over the five surge events discussed on this article. A negative delay indicates that the peak surge occurred later than at Tofino.}} 
{\begin{tabular}{|l |c | c|}
\hline 
Location & Mean surge amplitude (m) & Mean delay from Tofino (hrs) \\
\hline
Port Hardy & 0.491 & -2.67\\
Winter Harbour & 0.572 & 0.333 \\
Tofino  & 0.707 & -- \\
Bamfield & 0.658 & -0.5 \\
Neah Bay & 0.658 & -1.83 \\
Toke Point & 1.17 & -2.83 \\
\hline
\end{tabular}}
\label{tab:pacificsurge}
\end{table}

@@

The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary. The barotropic velocity component of the remote forcing was set to zero. At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).  This scheme is not part of core NEMO 3.4 (the version used here).  Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain. {\color{red} No boundary conditions were set on the baroclinic velocities at the northern boundary.}

At coastal boundaries, a partial slip boundary condition, an approximation to no slip, is used. Partial slip allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions. A lateral eddy viscosity of 20 m$^2$ s$^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m$^2$ s$^{-1}$ is used.  Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$. Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ m$^2$ s$^{-1}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively. Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.

@@

In general, the K$_1$ tidal harmonics are well-represented by the model with complex differences below 5 cm at all stations except Seymour Narrows (station 27). Seymour Narrows is a very constricted region in Johnstone Strait and is not well resolved by the model. As such, the M$_2$ amplitude and phase are also poorly reproduced here.  The model's M$_2$ amphidrome, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region. {\color{red} The model also represents the dip in M$_2$ amplitude around stations 25 and 26 reasonably well.} Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.  The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.  {\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised; the effect is particularly noticeable at the station closest to the open boundary (Port Renfrew, 1). } Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.

@@

Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$. Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range. {\color{red} As such, model water level predictions were corrected by accounting for this error. Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations and so the correction as calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (Sa, Ssa, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.  The difference between these two predictions is added to the model sea surface height as a correction. Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing, i.e.\, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy before the harmonic analysis is performed. The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals. All times are reported in UTC.

@@

The model's skill is reported in terms of statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
\begin{equation}
\bar{e} = \text{mean}\left(\left| M - O \right|\right),
\end{equation}
where $M$ and $O$ are the modelled and observed quantities of interest, is a common measure that estimates the average error in the model. Next, a measure comparing the variance of the error to the variance of the observations defined as \citep{thompson2003prediction},
\begin{equation}
\gamma^2 = \frac{\text{var}\left(M-O\right)}{\text{var}\left(O\right)},
\end{equation}
is also calculated. Smaller values of $\gamma^2$ indicate better model skill. This quantity was used by \citet{bernier2006predicting} and \citet{bernier2010tide} in storm surge predictions in Eastern Canada and is calculated here as a point of comparison. A third measure, called the Willmott skill score \citep{willmott1982some}, is also used to assess overall model skill in predicting the water level and residual. This quantity is defined as
\begin{equation}
WS = 1 - \frac{\sum_{i=1}^N \left(M_i - O_i\right)^2}{\sum_{i=1}^N \left(|M_i'| + |O_i'|\right)^2},
\end{equation} 
where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations. This score is bounded by zero and one and values close to one indicate good model skill. Since the Willmott skill score is nondimensional, it is an appropriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals. Since the total water level has a larger variability than the residuals, the $\gamma^2$ measure should be much lower for the total water levels; hence, the Willmott skill score provides an additional point of comparison. These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.  {\color{red} Observed and modelled maximum water levels and surge values as well as differences in their timing are reported in Table \ref{tab:peak}.}

@@

{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and briefly examine the effects of stratification. When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm. The other stations were not strongly affected. This was also observed in another hindcast (Dec 2006). The effect of stratification was examined by considering a case with an initially horizontally uniform stratification and comparing to the fully stratified case. The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period. Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic model as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}. We plan to investigate stratified effects in more detail in a future study.}

@@

A simple steady state argument balancing the wind stress with the pressure gradient can be used to calculate the slope of the sea surface in a body of water with a constant depth $D$ \citep{pugh2004changing}: 
\begin{equation}
\text{slope} = \frac{C\rho_{air}U^2}{g\rho D},
\end{equation}
where $C$ is the coefficient of friction, $\rho_{air}$ is the density of air, $g$ is the acceleration due to gravity, $U$ is the wind speed, and $\rho$ is the density of water. Assuming  $C=10^{-3}$, $\rho_{air}=1.23$ kg m$^{-3}$, $\rho= 1035$  kg m$^{-3}$, wind speed $U=20$ m s$^{-1}$, and depth $D=100$ m, the slope of the sea surface is approximately $5\times10^{-7}$. At Point Atkinson, a westerly wind can act over a small distance of approximately 50 km, giving a sea surface elevation on the order of centimetres in agreement with the model findings. Since the Strait of Georgia is not a constant depth and is up to 400 m deep, this calculation is only an approximation. Choosing a larger depth would result in an even smaller estimate for the sea surface elevation caused by wind stresses. Further, northwesterly winds would induce the largest elevations at Point Atkinson due to the large distance over which the winds can act.

%local pressure
Next, the surge amplitude decreases by a few centimetres at all four locations if local atmospheric pressure is not included in the surface forcing. This drop is smaller than the IB sea level change calculated by equation (\ref{eq:inverse}), approximately 1 cm rise in sea level associated with 1 hPa drop in atmospheric pressure. In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level. This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to one third of the total surge amplitude. In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing. Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response. {\color{red} It is the atmopsheric pressure gradient that drives changes in the sea surface height. Since our domain is relatively small, so is the atmospheric pressure gradient, and hence, we see only a small contribution from local atmospheric pressure. }

@@

Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well. All Willmott skill scores rank above {\color{red}0.89} and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}). The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals. Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases. This could be due to a number of factors. First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast. Second, the anomaly forcing at the open boundaries could be too low. The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain). {\color{red} Since the storm surge is less sensitive to wind forcing than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}

@@

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.  {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the incoming surge wave could be partially reflected from the mainland coast in the Strait of Georgia. Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case. However, this is still less than 10 \% of the total surge amplitude. }

@@

{\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.  Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and {\color{red} the direction towards which the wind is heading,} at the closest grid point to the longitude/latitude of Point Atkinson, Sandheads, and the NOAA buoy, is used for comparison. Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations. At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012. At Sandheads, both the low resolution and high resolution models have good agreement with observations. The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water. Although not shown here, both models have a similar sea level atmospheric pressure at these locations. Note that data assimilation is used in the parent model of the nested HRDPS product.

\begin{figure}
\centering
\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
\caption{{\color{red}Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (top), Sandheads (middle) and NOAA Buoy 46087 (bottom). Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey. Observations from Sandheads and Point Atkinson are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.  Observations at the NOAA Buoy were downloaded from: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}. Wind direction represents the direction that the wind is headed.} }
\label{fig:dec2012_weather}
\end{figure}
